---
title: "Steroid targeted assay analyses"
output: html_document
date: "2022-10-23"
---

### Load libraries
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
library(ggplot2)
library(janitor)
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(patchwork)
library(dplyr)

#set theme for plots in this markdown
theme_set(theme_minimal())
```

```{r setup, include=FALSE}
#set preference for markdown chunks
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### Load data and merge with metadata
Load metadata file ("metadata.csv") containing patients info including Subject ID, gestational age at sampling, outcome info, ethnicity, BMI etc. 
Load steroids levels files ("hormones_values_final.csv") containing final values calculated with adjustment for sample weight. A detailed explanation of the calculations done is in the file named "hormones_calculations.csv".
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Load metadata
metadata<- read.csv(here("metadata.csv"))
metadata<- clean_names(metadata)

#Load steroids values
hormones<- read.csv((here("hormones_values_final.csv")))
hormones<- clean_names(hormones)

#Merge by sample_id
data <- merge(metadata, hormones, by="sample_id")

#Recode Miscarriage (n=1) as preterm
data<- data %>% mutate (preterm_v_term_1 = case_when(preterm_v_term == "Preterm" ~ "Preterm",
                                                     preterm_v_term == "Miscarriage" ~ "Preterm",
                                                     preterm_v_term == "Term" ~ "Term")) %>%
                                           relocate(preterm_v_term_1, .after = preterm_v_term)

```

### Basic summary of statistics - all samples included
Summary of basic statistics for progesterone, cortisol and androstenedione levels in all samples. Results saved in "summary_levels_P_C_A.csv". 
```{r}

#Summary statistics progesterone
summary_P<- data %>%
  summarise(
    count = n(),
    min = min(data$progesterone_ng_g, na.rm = TRUE),
    median = median(data$progesterone_ng_g, na.rm = TRUE),
    max = max(data$progesterone_ng_g, na.rm = TRUE),
    sd = sd(data$progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(data$progesterone_ng_g, na.rm = TRUE))

#Summary statistics cortisol
summary_C<- data %>%
  summarise(
    count = n(),
    min = min(data$cortisol_ng_g, na.rm = TRUE),
    median = median(data$cortisol_ng_g, na.rm = TRUE),
    max = max(data$cortisol_ng_g, na.rm = TRUE),
    sd = sd(data$cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(data$cortisol_ng_g, na.rm = TRUE))

#Summary statistics androstenedione
summary_A<- data %>%
  summarise(
    count = n(),
    min = min(data$androstenedione_ng_g, na.rm = TRUE),
    median = median(data$androstenedione_ng_g, na.rm = TRUE),
    max = max(data$androstenedione_ng_g, na.rm = TRUE),
    sd = sd(data$androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(data$androstenedione_ng_g, na.rm = TRUE))

#save results in csv file
#write.csv(rbind(summary_P, summary_C, summary_A), here("images", "summary_levels_P_C_A.csv"))
summary_P
summary_C
summary_A
```

### Box plots - all samples including
Box plots showing levels of the steroids in all samples available. Image with 3 plots side by side also shown.
```{r}

#box plot of progesterone levels 
P <- ggplot(data, aes(x = 1, y = data$progesterone_ng_g)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Progesterone") + ylab("ng/g") +
     theme(legend.position = "none", text = element_text(size=20)) + 
     ylim(0, 20)
#ggsave(here("images", "progesterone_levels.png"), P, width = 3, height = 8)

#box plot of cortisol levels 
C <- ggplot(data, aes(x = 1, y = data$cortisol_ng_g)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Cortisol") + ylab("ng/g") +
     theme(legend.position = "none", text = element_text(size=20)) +
     ylim(0, 20)
#ggsave(here("images", "cortisol_levels.png"), C, width = 3, height = 8)

#box plot of androstenedione levels 
A <- ggplot(data, aes(x = 1, y = data$androstenedione_ng_g)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Androstenedione") + ylab("ng/g") +
     theme(legend.position = "none", text = element_text(size=20)) + 
      ylim(0, 20)
#ggsave(here("images", "androstenedione_levels.png"), A, width = 3, height = 8)


#image with 3 plots together
P_C_A_plot<-P + C+ A & ylab(NULL)
P_C_A_plot<-wrap_elements(P_C_A_plot) +
  labs(tag = "ng/g") +
  theme(plot.tag = element_text(size = 20, angle = 90),
  plot.tag.position = "left")

#ggsave(here("images", "prog_cort_and_levels.png"), P_C_A_plot, width = 8, height = 8)
P_C_A_plot
```

### Basic summary of statistics - only low-risk samples
Summary of basic statistics for progesterone, cortisol and androstenedione levels in low risk samples.This is done to have levels of main steroids in healthy pregnancies. Results saved in "summary_levels_P_C_A_controls.csv". 
```{r, echo=TRUE, message=FALSE, warning=FALSE}

#filter out high risk pregnancy and term pregnancy
data_low_risk<- data %>% filter(data$risk_group == "control" & data$preterm_v_term_1 == "Term")


summary_P<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    median = median(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    max = max(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    sd = sd(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(data_low_risk$progesterone_ng_g, na.rm = TRUE))

summary_C<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    median = median(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    max = max(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    sd = sd(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(data_low_risk$cortisol_ng_g, na.rm = TRUE))

summary_A<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    median = median(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    max = max(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    sd = sd(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(data_low_risk$androstenedione_ng_g, na.rm = TRUE))

#write.csv(rbind(summary_P, summary_C, summary_A), here("images","summary_levels_P_C_A_controls.csv"))

summary_P
summary_C
summary_A
```

### Box plots - low risk samples
Box plots showing levels of the steroids low risk samples. Image with 3 plots side by side also shown.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#box plot of progesterone levels 
P <- ggplot(data_low_risk, aes(x = 1, y = data_low_risk$progesterone_ng_g)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Progesterone") + ylab("ng/g") +
     theme(legend.position = "none", text = element_text(size=20)) +
     ylim(0, 25)
#ggsave(here("images", "progesterone_levels_LowRisk.png"), P, width = 3, height = 8)

#box plot of cortisol levels 
C <- ggplot(data_low_risk, aes(x = 1, y = data_low_risk$cortisol_ng_g)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Cortisol") + ylab("ng/g") +
     theme(legend.position = "none", text = element_text(size=20)) +
     ylim(0, 25)
#(here("images", "cortisol_levels_LowRisk.png"), C, width = 3, height = 8)

#box plot of androstenedione levels 
A <- ggplot(data_low_risk, aes(x = 1, y = data_low_risk$androstenedione_ng_g)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Androstenedione") + ylab("ng/g") +
     theme(legend.position = "none", text = element_text(size=20)) +
     ylim(0, 25)
#ggsave(here("images", "androstenedione_levels_LowRisk.png"), A, width = 3, height = 8)


#image with 3 plots together
P_C_A_plot<-P + C+ A & ylab(NULL)
P_C_A_plot<-wrap_elements(P_C_A_plot) +
  labs(tag = "ng/g") +
  theme(plot.tag = element_text(size = 20, angle = 90),
  plot.tag.position = "left")

#ggsave(here("images", "prog_cort_and_levels_LowRisk.png"), P_C_A_plot, width = 8, height = 8)
P_C_A_plot
```

***

## Analysis of steroids relationship with gestational age
### Progesterone
Linear model progesterone ~ gestational age
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#linear model prog vs gest age
model<- lm(data$progesterone_ng_g ~ gestational_age_final , data= data)
s<- summary(model)

R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)

# Create label
label_text <- paste0("R² = ", R2,
                     "\nF(", df1, ", ", df2, ") = ", Fstat,
                     "\np = ", pval)

#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot coloured according to risk factor annotated with stats
prog_scatterplot <- ggplot(data, aes(x = gestational_age_final, y = progesterone_ng_g)) +
  geom_point(aes(colour = risk_group), size = 4) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction)) +
  annotate("text", 
           x = 13, y = max(data$progesterone_ng_g, na.rm = TRUE)*0.95, 
           label = label_text, hjust = 0, size = 5) +
  labs(colour = "Outcome") +
  xlab("Gestational Age (weeks)") +
  ylab("Progesterone (ng/g)") +
  theme(text = element_text(size = 15)) +
  scale_colour_manual(values = c("control" = "#00BFC4", "high" = "#F8766D"),  labels = c("control" = "Uncomplicated", "high" = "Adverse"))

#ggsave(here("images_reviewed", "prog_scatterplot_coloured.png"), prog_scatterplot, width = 8, height = 8)
prog_scatterplot
```

### Cortisol
Linear model cortisol ~ gestational age
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#linear model prog vs gest age
model_cort<- lm(cortisol_ng_g ~ gestational_age_final , data= data)
sc<- summary(model_cort)

R2 <- round(sc$r.squared, 3)
Fstat <- round(sc$fstatistic[1], 3)
df1 <- sc$fstatistic[2]
df2 <- sc$fstatistic[3]
pval <- signif(pf(sc$fstatistic[1], df1, df2, lower.tail = FALSE), 3)

# Create label
label_text <- paste0("R² = ", R2,
                     "\nF(", df1, ", ", df2, ") = ", Fstat,
                     "\np = ", pval)


#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model_cort, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot coloured according to risk factor annotated with stats
cort_scatterplot <- ggplot(data, aes(x = gestational_age_final, y = cortisol_ng_g)) +
  geom_point(aes(colour = risk_group), size = 4) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction)) +
  annotate("text", 
           x = 13, y = max(data$cortisol_ng_g, na.rm = TRUE)*0.95, 
           label = label_text, hjust = 0, size = 5) +
  labs(colour = "Outcome") +
  xlab("Gestational Age (weeks)") +
  ylab("Cortisol (ng/g)") +
  theme(text = element_text(size = 15)) +
  scale_colour_manual(values = c("control" = "#00BFC4", "high" = "#F8766D"), labels = c("control" = "Uncomplicated", "high" = "Adverse"))

#ggsave(here("images_reviewed", "cort_scatterplot_coloured.png"), cort_scatterplot, width = 8, height = 8)
cort_scatterplot
```

### Androstenedione
Linear model androstenedione ~ gestational age
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#linear model prog vs gest age
model_and<- lm(androstenedione_ng_g ~ gestational_age_final , data= data)
sa<- summary(model_and)

R2 <- round(sa$r.squared, 3)
Fstat <- round(sa$fstatistic[1], 3)
df1 <- sa$fstatistic[2]
df2 <- sa$fstatistic[3]
pval <- signif(pf(sa$fstatistic[1], df1, df2, lower.tail = FALSE), 3)

# Create label
label_text <- paste0("R² = ", R2,
                     "\nF(", df1, ", ", df2, ") = ", Fstat,
                     "\np = ", pval)


#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model_and, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot coloured according to risk factor
androst_scatterplot<- ggplot(data, aes(x = gestational_age_final, y = androstenedione_ng_g)) +
  geom_point(aes(colour = risk_group), size = 4) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction)) +
  annotate("text", 
           x = 13, y = max(data$androstenedione_ng_g, na.rm = TRUE)*0.95, 
           label = label_text, hjust = 0, size = 5) +
  labs(colour = "Outcome") +
  xlab("Gestational Age (weeks)") +
  ylab("Androstenedione (ng/g)") +
  theme(text = element_text(size = 15)) +
  scale_colour_manual(values = c("control" = "#00BFC4", "high" = "#F8766D"), labels = c("control" = "Uncomplicated", "high" = "Adverse"))



#ggsave(here("images_reviewed", "androst_scatterplot_coloured.png"), androst_scatterplot, width = 8, height = 8)
androst_scatterplot
```





### Figure with the 3 scatterplots
```{r, echo=TRUE, message=FALSE, warning=FALSE}
scatterplots<- prog_scatterplot+ cort_scatterplot+ androst_scatterplot + plot_layout(guides = 'collect')
ggsave(here("images_reviewed", "scatterplots_P_C_A.png"), scatterplots, width = 18, height = 6)
```


***

## Relationship with gestational age in longitudinal samples
Investigate what happens to hormones levels for individuals were we have longitudinal samples (n=5)

```{r}
#filter only longitudinal data
data_paired<- data %>% filter(data$longitudinal =="yes")
```

### Progesterone
```{r}
#progesterone individuals levels
individuals_prog<- data_paired %>%
  ggplot(aes(gestational_age_final,progesterone_ng_g, color=timepoint)) +
  geom_point(aes(fill=timepoint),size=5) +
  geom_line(aes(group = cons_v2),color="grey")+ xlab("Gestational Age") + ylab("Progesterone ng/g") + theme(text = element_text(size = 20), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 10))+ facet_grid(. ~ cons_v2) + scale_colour_manual(values = c("A" = "#a5e35d", "B" = "#388E3C"))
#ggsave(here("images_reviewed", "individuals_prog_coloured.png"), individuals_prog, width = 8, height = 8)

#statistical summary
group_by(data_paired, timepoint) %>%
  summarise(
    count = n(),
    median = median(progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(progesterone_ng_g, na.rm = TRUE))

#box plot
prog_timepoint_A_B<- ggplot(data_paired, aes(x = data_paired$timepoint, y = as.numeric(data_paired$progesterone_ng_g)))+ geom_boxplot(aes(fill=data_paired$timepoint), alpha = 0.8, outlier.shape = NA)+ xlab("Timepoint")+ylab("Progesterone")+theme(legend.position = "none")+ theme(text = element_text(size = 30))  + scale_fill_manual(values = c("A" = "#a5e35d", "B" = "#388E3C"))+ ylim(0, 20)
#ggsave(here("images_reviewed", "prog_timepoint_A_B_coloured.png"), prog_timepoint_A_B, width = 8, height = 8)

#t-test
numeric_timepoint <- ifelse( data_paired$timepoint == "A", 1, 2)
result <- t.test(data_paired$progesterone_ng_g, numeric_timepoint, paired = TRUE)
result$p.value

individuals_prog
prog_timepoint_A_B
``` 

### Cortisol
```{r}
#Cortisol individual levels
individuals_cort<- data_paired %>%
  ggplot(aes(gestational_age_final,cortisol_ng_g, color=timepoint)) +
  geom_point(aes(fill=timepoint),size=5) +
  geom_line(aes(group = cons_v2),color="grey")+ xlab("Gestational Age") + ylab("Cortisol ng/g") + theme(text = element_text(size = 20), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 10))+ facet_grid(. ~ cons_v2)  + scale_colour_manual(values = c("A" = "#a5e35d", "B" = "#388E3C"))
#ggsave(here("images_reviewed", "individuals_cort_coloured.png"), individuals_cort, width = 8, height = 8)

#statistical summary
group_by(data_paired, timepoint) %>%
  summarise(
    count = n(),
    median = median(cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(cortisol_ng_g, na.rm = TRUE))

#box plot
cort_timepoint_A_B<- ggplot(data_paired, aes(x = data_paired$timepoint, y = as.numeric(data_paired$cortisol_ng_g)))+ geom_boxplot(aes(fill=data_paired$timepoint), alpha = 0.8, outlier.shape = NA)+ xlab("Timepoint")+ylab("Cortisol")+theme(legend.position = "none")+ theme(text = element_text(size = 30)) + scale_fill_manual(values = c("A" = "#a5e35d", "B" = "#388E3C"))
#ggsave(here("images_reviewed", "cort_timepoint_A_B_coloured.png"), cort_timepoint_A_B, width = 8, height = 8)

#t-test
numeric_timepoint <- ifelse( data_paired$timepoint == "A", 1, 2)
result <- t.test(data_paired$cortisol_ng_g, numeric_timepoint, paired = TRUE)
result$p.value

individuals_cort
cort_timepoint_A_B
``` 

### Androstenedione
```{r}
#Cortisol individual levels
individuals_andr<- data_paired %>%
  ggplot(aes(gestational_age_final,androstenedione_ng_g, color=timepoint)) +
  geom_point(aes(fill=timepoint),size=5) +
  geom_line(aes(group = cons_v2),color="grey")+ xlab("Gestational Age") + ylab("Androstenedione ng/g") + theme(text = element_text(size = 20), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 10))+ facet_grid(. ~ cons_v2)  + scale_colour_manual(values = c("A" = "#a5e35d", "B" = "#388E3C"))
#ggsave(here("images_reviewed", "individuals_andr_coloured.png"), individuals_andr, width = 8, height = 8)

#statistical summary
group_by(data_paired, timepoint) %>%
  summarise(
    count = n(),
    median = median(androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(androstenedione_ng_g, na.rm = TRUE))

#box plot
andr_timepoint_A_B<- ggplot(data_paired, aes(x = data_paired$timepoint, y = as.numeric(data_paired$androstenedione_ng_g)))+ geom_boxplot(aes(fill=data_paired$timepoint), alpha = 0.8, outlier.shape = NA)+ xlab("Timepoint")+ylab("Androstenedione")+theme(legend.position = "none")+ theme(text = element_text(size = 30)) + scale_fill_manual(values = c("A" = "#a5e35d", "B" = "#388E3C"))
#ggsave(here("images_reviewed", "andr_timepoint_A_B_coloured.png"), andr_timepoint_A_B, width = 8, height = 8)

#t-test
numeric_timepoint <- ifelse(data_paired$timepoint == "A", 1, 2)
result <- t.test(data_paired$androstenedione_ng_g, numeric_timepoint, paired = TRUE)
result$p.value

individuals_andr
andr_timepoint_A_B
``` 

***

***

# Regression model of cervical lenght and steroids levels


#Progesterone plot
```{r}
data$cl_visit <- as.numeric(as.character(data$cl_visit))

# Remove missing value from miscarriage patient
data_clean <- data[is.finite(data$cl_visit) & is.finite(data$progesterone_ng_g), ]

# Fit linear model
model_cl <- lm(progesterone_ng_g ~ cl_visit, data = data_clean)

# Summary
s <- summary(model_cl)

# Extract stats
R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)

label_text <- paste0("R² = ", R2,
                     "\nF(", df1, ", ", df2, ") = ", Fstat,
                     "\np = ", pval)

# Prediction frame
fake_cl <- data.frame(cl_visit = seq(min(data_clean$cl_visit), 
                                     max(data_clean$cl_visit), 
                                     length.out = 100))
prediction <- predict(model_cl, newdata = fake_cl)
predicted_frame <- cbind(fake_cl, prediction)

# Plot
prog_cerv<- ggplot(data_clean, aes(x = cl_visit, y = progesterone_ng_g)) +
  geom_point(aes(colour = risk_group), size = 4) +
  geom_line(data = predicted_frame, aes(x = cl_visit, y = prediction)) +
  annotate("text", 
           x = min(data_clean$cl_visit) + 0.2,
           y = max(data_clean$progesterone_ng_g)* 0.95, 
           label = label_text, 
           hjust = 0, size = 6) +
  labs(colour = "Outcome") +
  xlab("Cervical Length (mm)") +
  ylab("Progesterone (ng/g)") +
  theme(text = element_text(size = 15)) +
  scale_colour_manual(values = c("control" = "#00BFC4", "high" = "#F8766D"), labels = c("control" = "Uncomplicated", "high" = "Adverse"))
#ggsave(here("images_reviewed", "prog_cerv.png"), prog_cerv, width = 6, height = 6)
```

# Cortisol plot

```{r}
model_cort <- lm(cortisol_ng_g ~ cl_visit, data = data_clean)
s <- summary(model_cort)

R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)

label_text <- paste0("R² = ", R2,
                     "\nF(", df1, ", ", df2, ") = ", Fstat,
                     "\np = ", pval)

fake_cl <- data.frame(cl_visit = seq(min(data_clean$cl_visit), max(data_clean$cl_visit), length.out = 100))
prediction <- predict(model_cort, newdata = fake_cl)
predicted_frame <- cbind(fake_cl, prediction)

cortisol_cerv<- ggplot(data_clean, aes(x = cl_visit, y = cortisol_ng_g)) +
  geom_point(aes(colour = risk_group), size = 4) +
  geom_line(data = predicted_frame, aes(x = cl_visit, y = prediction)) +
  annotate("text", x = min(data_clean$cl_visit) + 0.2, y = max(data_clean$cortisol_ng_g)* 0.95, 
           label = label_text, hjust = 0, size = 6) +
  labs(colour = "Outcome") +
  xlab("Cervical Length (mm)") +
  ylab("Cortisol (ng/g)") +
  theme(text = element_text(size = 15)) +
  scale_colour_manual(values = c("control" = "#00BFC4", "high" = "#F8766D"), labels = c("control" = "Uncomplicated", "high" = "Adverse"))
#ggsave(here("images_reviewed", "cortisol_cerv.png"), cortisol_cerv, width = 6, height = 6)

```
#Androstenedione

```{r}
model_and <- lm(androstenedione_ng_g ~ cl_visit, data = data_clean)
s <- summary(model_and)

R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)

label_text <- paste0("R² = ", R2,
                     "\nF(", df1, ", ", df2, ") = ", Fstat,
                     "\np = ", pval)

fake_cl <- data.frame(cl_visit = seq(min(data_clean$cl_visit), max(data_clean$cl_visit), length.out = 100))
prediction <- predict(model_and, newdata = fake_cl)
predicted_frame <- cbind(fake_cl, prediction)

andr_cerv<- ggplot(data_clean, aes(x = cl_visit, y = androstenedione_ng_g)) +
  geom_point(aes(colour = risk_group), size = 4) +
  geom_line(data = predicted_frame, aes(x = cl_visit, y = prediction)) +
  annotate("text", x = min(data_clean$cl_visit) + 0.2, y = max(data_clean$androstenedione_ng_g)* 0.95, 
           label = label_text, hjust = 0, size = 6) +
  labs(colour = "Outcome") +
  xlab("Cervical Length (mm)") +
  ylab("Androstenedione (ng/g)") +
  theme(text = element_text(size = 15)) +
  scale_colour_manual(values = c("control" = "#00BFC4", "high" = "#F8766D"), labels = c("control" = "Uncomplicated", "high" = "Adverse"))
#ggsave(here("images_reviewed", "andr_cerv.png"), andr_cerv, width = 6, height = 6)

```




### Figure with the 3 scatterplots
```{r, echo=TRUE, message=FALSE, warning=FALSE}
scatterplots<- prog_cerv + cortisol_cerv + andr_cerv + plot_layout(guides = 'collect')
ggsave(here("images_reviewed", "scatterplots_P_C_A_cerv_lenght.png"), scatterplots, width = 18, height = 6)
```






## Difference in steroids levels in term vs preterm
Analyses performed considering only one sample per patient (select timepoint B for longitudinal ones to balance gestational age of preterm group).

```{r}
#filter data
data_lo<- data %>% filter(data$longitudinal =="yes" & data$timepoint == "B")
data_nolo<- data %>% filter(data$longitudinal =="no")

data_oneeach<- rbind(data_lo, data_nolo)
```

### Progesterone

```{r}
#box plot
boxplot_outcome_prog<- ggplot(data_oneeach, aes(x = data_oneeach$preterm_v_term_1, y = as.numeric(data_oneeach$progesterone_ng_g)))+                 geom_boxplot(aes(fill=data_oneeach$preterm_v_term_1), alpha = 0.8, outlier.shape = NA)+ xlab("Outcome")+ylab("Progesterone")+ theme(legend.position = "none", text = element_text(size = 15))+ scale_fill_manual(values = c("Term"= "#1cd41c", "Preterm"= "#e34a0e"))
#ggsave(here("images_reviewed", "boxplot_outcome_prog_coloured.png"), boxplot_outcome_prog, width = 6, height = 6)

#statistical summary
group_by(data_oneeach, preterm_v_term_1) %>%
  summarise(
    count = n(),
    median = median(progesterone_ng_g, na.rm = TRUE),
    mean=mean(progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(progesterone_ng_g, na.rm = TRUE),
    SD = sd(progesterone_ng_g))

#wilcoxon test
res <- wilcox.test(progesterone_ng_g ~ preterm_v_term_1, data = data_oneeach,
                   exact = FALSE)
res
boxplot_outcome_prog
```


### Cortisol

```{r}
#box plot
boxplot_outcome_cort<- ggplot(data_oneeach, aes(x = data_oneeach$preterm_v_term_1, y = as.numeric(data_oneeach$cortisol_ng_g)))+                 geom_boxplot(aes(fill=data_oneeach$preterm_v_term_1), alpha = 0.8, outlier.shape = NA)+ xlab("Outcome")+ylab("Cortisol")+ theme(legend.position = "none", text = element_text(size = 15))+ scale_fill_manual(values = c("Term"= "#1cd41c", "Preterm"= "#e34a0e"))
#ggsave(here("images_reviewed", "boxplot_outcome_cort_coloured.png"), boxplot_outcome_cort, width = 6, height = 6)

#statistical summary
group_by(data_oneeach, preterm_v_term_1) %>%
  summarise(
    count = n(),
    median = median(cortisol_ng_g, na.rm = TRUE),
    mean=mean(cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(cortisol_ng_g, na.rm = TRUE),
    SD = sd(cortisol_ng_g))

#wilcoxon test
res <- wilcox.test(cortisol_ng_g ~ preterm_v_term_1, data = data_oneeach,
                   exact = FALSE)
res
boxplot_outcome_cort
```

### Androstenedione
```{r}
#boxplot
boxplot_outcome_andr<- ggplot(data_oneeach, aes(x = data_oneeach$preterm_v_term_1, y = as.numeric(data_oneeach$androstenedione_ng_g)))+                 geom_boxplot(aes(fill=data_oneeach$preterm_v_term_1), alpha = 0.8, outlier.shape = NA)+ xlab("Outcome")+ylab("Androstenedione")+ theme(legend.position = "none", text = element_text(size = 15))+ scale_fill_manual(values = c("Term"= "#1cd41c", "Preterm"= "#e34a0e"))
ggsave(here("images_reviewed", "boxplot_outcome_andr_coloured.png"), boxplot_outcome_andr, width = 6, height = 6)

#statistical summary
group_by(data_oneeach, preterm_v_term_1) %>%
  summarise(
    count = n(),
    median = median(androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(androstenedione_ng_g, na.rm = TRUE),
    SD = sd(androstenedione_ng_g))

#wilcoxon test
res <- wilcox.test(androstenedione_ng_g ~ preterm_v_term_1, data = data_oneeach,
                   exact = FALSE)
res
boxplot_outcome_andr
```

### save image with 3 boxplots
```{r}
boxplot_outcome_all<- boxplot_outcome_prog + boxplot_outcome_cort + boxplot_outcome_andr + plot_layout(guides = 'collect')
ggsave(here("images_reviewed", "boxplot_outcome_P_C_A.png"), boxplot_outcome_all, width = 18, height = 6)
```


```{r}
# Combine rows
combined_figure <- scatterplots / boxplot_outcome_all +
    plot_layout(heights = c(1, 1.2))  # adjust spacing

# Save
ggsave(here("images_reviewed", "combined_scatter_boxplots_panelAB.png"), 
       combined_figure, width = 18, height = 12)
```



***



```{r}
# Convert weeks + days into decimal weeks
data$gestational_age_delivery_final <- data$gestation_at_delivery + data$gestation_at_delivery_days / 7

# Check
head(data[, c("gestation_at_delivery", "gestation_at_delivery_days", "gestational_age_delivery_final")])


```

```{r}

# Progesterone
model_prog_gad <- lm(progesterone_ng_g ~ gestational_age_delivery_final, data = data)
summary(model_prog_gad)

# Cortisol
model_cort_gad <- lm(cortisol_ng_g ~ gestational_age_delivery_final, data = data)
summary(model_cort_gad)

# Androstenedione
model_and_gad <- lm(androstenedione_ng_g ~ gestational_age_delivery_final, data = data)
summary(model_and_gad)

```
```{r}
# Function to extract R², F-statistic, p-value
get_stats <- function(model){
  s <- summary(model)
  R2 <- round(s$r.squared, 3)
  Fstat <- round(s$fstatistic[1], 3)
  df1 <- s$fstatistic[2]
  df2 <- s$fstatistic[3]
  pval <- signif(pf(Fstat, df1, df2, lower.tail = FALSE), 3)
  return(list(R2 = R2, Fstat = Fstat, df1 = df1, df2 = df2, pval = pval))
}

stats_prog <- get_stats(model_prog_gad)
stats_cort <- get_stats(model_cort_gad)
stats_and <- get_stats(model_and_gad)
```



```{r}


# Function to create plot
plot_steroid <- function(data, yvar, model, stats, ylab){
  
  # Prediction frame
  fake_x <- data.frame(gestational_age_delivery_final = seq(min(data$gestational_age_delivery_final, na.rm = TRUE),
                                                      max(data$gestational_age_delivery_final, na.rm = TRUE),
                                                      length.out = 100))
  prediction <- predict(model, newdata = fake_x)
  predicted_frame <- cbind(fake_x, prediction)
  
  # Label text
  label_text <- paste0("R² = ", stats$R2, 
                       "\np = ", stats$pval)
  
  # Plot
  ggplot(data, aes(x = gestational_age_delivery_final, y = !!sym(yvar))) +
    geom_point(aes(colour = risk_group), size = 3) +
    geom_line(data = predicted_frame, aes(x = gestational_age_delivery_final, y = prediction), color = "black", linetype= "dashed") +
    annotate("text", x = min(data$gestational_age_delivery_final, na.rm = TRUE),
             y = max(data[[yvar]], na.rm = TRUE),
             label = label_text, hjust = 0, vjust = 1, size = 5) +
    scale_colour_manual(values = c("control" = "#00BFC4" , "high" = "#F8766D" ), labels = c("control" = "Uncomplicated", "high" = "Adverse"))+
    labs(x = "Gestational Age at Delivery (weeks)", y = ylab, colour = "Outcome") +
    theme_minimal() +
  theme(axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14))
}

# Create plots
prog_plot <- plot_steroid(data, "progesterone_ng_g", model_prog_gad, stats_prog, "Progesterone (ng/g)")
cort_plot <- plot_steroid(data, "cortisol_ng_g", model_cort_gad, stats_cort, "Cortisol (ng/g)")
and_plot <- plot_steroid(data, "androstenedione_ng_g", model_and_gad, stats_and, "Androstenedione (ng/g)")

# Display example
prog_plot
cort_plot
and_plot


# Combine plots in a single row
combined_plot <- prog_plot + cort_plot + and_plot + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom", legend.title = element_text(size=20), legend.text = element_text(size=20))  # shared legend

# Save
ggsave(here("images_reviewed", "steroids_vs_gest_age_delivery_noF.png"),
       combined_plot, width = 18, height = 6)


```




### check if a specific patient has anomalous levels of the steroids


```{r}
# Example for progesterone
patient_sample <- subset(data, sample_id == "2V1592_23+5")

patient_sample

controls <- subset(data, risk_group == "control")

range_prog_ctrl <- range(controls$progesterone_ng_g, na.rm = TRUE)
range_cort_ctrl <- range(controls$cortisol_ng_g, na.rm = TRUE)
range_andr_ctrl <- range(controls$androstenedione_ng_g, na.rm = TRUE)

range_prog_ctrl
range_cort_ctrl
range_andr_ctrl
# Check if within range
prog_in_range <- patient_sample$progesterone_ng_g >= range_prog_ctrl[1] & patient_sample$progesterone_ng_g <= range_prog_ctrl[2]
cort_in_range <- patient_sample$cortisol_ng_g >= range_cort_ctrl[1] & patient_sample$cortisol_ng_g <= range_cort_ctrl[2]
andr_in_range <- patient_sample$androstenedione_ng_g >= range_andr_ctrl[1] & patient_sample$androstenedione_ng_g <= range_andr_ctrl[2]

# Print results
cat("Progesterone within control range:", prog_in_range, "\n")
cat("Cortisol within control range:", cort_in_range, "\n")
cat("Androstenedione within control range:", andr_in_range, "\n")

```


### Multivariate Regression

```{r}

# Convert risk_group to numeric 0/1
data_clean$risk_bin <- ifelse(data_clean$risk_group == "high", 1, 0)

# Run logistic regression
model_multi <- glm(risk_bin ~ progesterone_ng_g + cortisol_ng_g + androstenedione_ng_g, 
                   data = data_clean, family = binomial)
summary(model_multi)
```




```{r}

#adjust for maternal confunders

model_multi_cov <- glm(risk_bin ~ progesterone_ng_g + cortisol_ng_g + androstenedione_ng_g +
                                      age + bmi + parity, data = data_clean, family = binomial)
summary(model_multi_cov)

```

### AUC for multivariate logistic model

```{r}

# Load packages
library(caret)
library(glmnet)
library(pROC)
library(dplyr)

# Prepare modeling dataset
data_model <- data_clean %>%
  select(risk_group,
         progesterone_ng_g,
         cortisol_ng_g,
         androstenedione_ng_g,
         age, bmi, parity) %>%
  na.omit()

# Ensure outcome is a factor with correct ordering
data_model$risk_group <- factor(
  data_model$risk_group,
  levels = c("control", "high")
)

# Set seed for reproducibility
set.seed(123)

# Define 3-fold cross-validation
ctrl <- trainControl(
  method = "cv",
  number = 3,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = TRUE
)

# Train logistic regression via caret (glmnet with no regularization)
model_cv <- train(
  risk_group ~ progesterone_ng_g +
               cortisol_ng_g +
               androstenedione_ng_g +
               age + bmi + parity,
  data = data_model,
  method = "glmnet",
  metric = "ROC",
  trControl = ctrl,
  tuneGrid = expand.grid(
    alpha = 0,
    lambda = 0
  ),
  family = "binomial"
)

# Extract cross-validated ROC AUC
cv_auc <- max(model_cv$results$ROC)
print(paste("3-fold cross-validated ROC AUC:", round(cv_auc, 3)))

# Plot ROC curve
roc_obj <- roc(
  response = model_cv$pred$obs,
  predictor = model_cv$pred$high
)

plot(
  roc_obj,
  main = paste("Cross-validated ROC AUC =", round(cv_auc, 3)),
  lwd = 3
)


png(here("images_reviewed", "cross_validated_ROC.png"),
width = 6, height = 6, units = "in", res = 300)

plot(
roc_obj,
main = paste("Cross-validated ROC AUC =", round(cv_auc, 3)),
lwd = 3
)

dev.off()

```







```{r}
pca_res <- prcomp(data_clean[, c("progesterone_ng_g","cortisol_ng_g","androstenedione_ng_g")], scale = TRUE)
scores <- as.data.frame(pca_res$x)
scores$risk_group <- data_clean$risk_group

library(ggplot2)
pca_plot<- ggplot(scores, aes(x = PC1, y = PC2, color = risk_group)) +
  geom_point(size = 4) + 
    scale_colour_manual(values = c("control" = "#00BFC4" , "high" = "#F8766D" ), labels = c("control" = "Uncomplicated", "high" = "Adverse"))+
  theme_minimal()+ labs(colour = "Outcome") +
  labs( x = paste0("PC1 (", round(var_exp[1], 1),"%)"),
    y = paste0("PC2 (", round(var_exp[2], 1), "%)")) +
  theme_minimal() +  theme(axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14), legend.title = element_text(size=18), legend.text = element_text(size=18))

ggsave(here("images_reviewed", "PCA_multivariate_adjusted.png"),
       pca_plot, width = 8, height = 6)


```



```{r}
# Proportion of variance explained
pca_var <- summary(pca_res)$importance[2, ]  # 2nd row = Proportion of Variance
pca_var

# Loadings
pca_res$rotation



```
```{r}
loadings <- as.data.frame(pca_res$rotation[, 1:2])
loadings$analyte <- rownames(pca_res$rotation)

# Apply renaming
loadings$analyte <- analyte_names[loadings$analyte]
loadings$analyte <- rownames(loadings)

loadings$analyte <- dplyr::recode(
  loadings$analyte,
  progesterone_ng_g    = "Progesterone",
  cortisol_ng_g        = "Cortisol",
  androstenedione_ng_g = "Androstenedione")

# Scaling factor
arrow_scale <- min(
  diff(range(scores$PC1)),
  diff(range(scores$PC2))) * 0.7

loadings$PC1 <- loadings$PC1 * arrow_scale
loadings$PC2 <- loadings$PC2 * arrow_scale


pca_plot_loadings <- ggplot(scores, aes(x = PC1, y = PC2, color = risk_group)) +
  geom_point(size = 4) +

  # Loadings arrows
  geom_segment(
    data = loadings,
    aes(x = 0, y = 0, xend = PC1, yend = PC2),
    arrow = arrow(length = unit(0.25, "cm")),
    inherit.aes = FALSE,
    color = "black" ) +

  # Loadings labels
  geom_text(
    data = loadings,
    aes(x = PC1, y = PC2, label = analyte),
    inherit.aes = FALSE,
    size = 5,
    vjust = -0.5) +

  scale_colour_manual(
    values = c("control" = "#00BFC4", "high" = "#F8766D"),
    labels = c("control" = "Uncomplicated", "high" = "Adverse")
  ) +

  labs(
    x = paste0("PC1 (", round(var_exp[1], 1), "%)"),
    y = paste0("PC2 (", round(var_exp[2], 1), "%)"), colour = "Outcome"
  ) +

  theme_minimal()+  theme(axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14), legend.title = element_text(size=18), legend.text = element_text(size=18))



ggsave(here("images_reviewed", "PCA_multivariate_adjusted.png"),
       pca_plot_loadings, width = 8, height = 6)





```






***


## Variation of CVF fluid weight collected with gestational age
Analyse if there is a significant difference in the weight of the CVF collected according to the gestational age

```{r}
#linear model weight vs gestational age
model<- lm(weight_g ~ gestational_age_final , data= data)
summary(model)

#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot
prog_scatterplot_all<- ggplot(data, aes(x=gestational_age_final, y=weight_g)) + geom_point(size = 4)+ geom_line(data=predicted_frame, aes(x=gestational_age_final, y=prediction))+ labs(colour= "Outcome")+ xlab("Gestational Age")+ylab("CVF weight")+ theme(text = element_text(size = 25)) 

ggsave(here("images", "weight_gestational_age.png"), prog_scatterplot_all, width = 8, height = 8)

prog_scatterplot_all
```


## Different weight for longitudinal sample timepoint A and B
```{r}

data_total_paired<- data %>% filter(data$longitudinal =="yes")

weight_timepoint_A_B<- ggplot(data_total_paired, aes(x = data_total_paired$timepoint, y = data_total_paired$weight_g))+ geom_boxplot(aes(fill=data_total_paired$timepoint), alpha = 0.4, outlier.shape = NA)+ xlab("Timepoint")+ylab("CVF weight")+theme(legend.position = "none")+ theme(text = element_text(size = 20)) 
ggsave(here("images", "weight_timepoint_A_B.png"), weight_timepoint_A_B, width = 8, height = 8)

#transform timepoint in numeric to do the analysis
numeric_timepoint <- ifelse(data_total_paired$timepoint == "A", 1, 2)
res_a <- wilcox.test(numeric_timepoint, data_total_paired$weight_g, data= data,
                   exact = FALSE, paired = TRUE, alternative = "two.sided")
res_a$p.value
```

***


***

## Analysis with values expressed in ng/ml
Repeat the basic analysis with the value recalculated in unit ng/ml
```{r}

hormones_ng_ml<- read.csv(here("hormones_ng_ml.csv"))

hormones_ng_ml<- clean_names(hormones_ng_ml)

data_ng_ml<-merge(metadata, hormones_ng_ml, by= "sample_id")

data_ng_ml<- data_ng_ml %>% mutate (preterm_v_term_1 = case_when(preterm_v_term == "Preterm" ~ "Preterm",
                                                     preterm_v_term == "Miscarriage" ~ "Preterm",
                                                     preterm_v_term == "Term" ~ "Term")) %>% relocate(preterm_v_term_1, .after = preterm_v_term)
```



### measurment of steroids only in low risk samples
```{r}

data_low_risk<- data_ng_ml %>% filter(data_ng_ml$risk_group == "control" & data_ng_ml$preterm_v_term_1 == "Term")

summary_P<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    median = median(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    max = max(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    sd = sd(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    IQR = IQR(data_low_risk$progesterone_ng_ml, na.rm = TRUE))

summary_C<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    median = median(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    max = max(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    sd = sd(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    IQR = IQR(data_low_risk$cortisol_ng_ml, na.rm = TRUE))

summary_A<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    median = median(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    max = max(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    sd = sd(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    IQR = IQR(data_low_risk$androstenedione_ng_ml, na.rm = TRUE))

write.csv(rbind(summary_P, summary_C, summary_A), here("images","summary_levels_NG_ML_controls.csv"))



P <- ggplot(data_low_risk, aes(x = 1, y = data_low_risk$progesterone_ng_ml)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Progesterone") + ylab("ng/ml") +
     theme(legend.position = "none") +
     ylim(0, 10)
ggsave(here("images", "progesterone_levels_ng_ml.png"), P, width = 3, height = 8)

C <- ggplot(data_low_risk, aes(x = 1, y = data_low_risk$cortisol_ng_ml)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Cortisol") + ylab("ng/ml") +
     theme(legend.position = "none") +
     ylim(0, 10)
ggsave(here("images", "cortisol_levels_ng_ml.png"), C, width = 3, height = 8)

A <- ggplot(data_low_risk, aes(x = 1, y = data_low_risk$androstenedione_ng_ml)) +
     geom_boxplot(fill = "#4682b4") +
     geom_jitter(width = 0.2, height = 0, color = "black", alpha = 0.5) +
     scale_x_continuous(breaks = NULL) +
     xlab("Androstenedione") + ylab("ng/ml") +
     theme(legend.position = "none") +
     ylim(0, 10)
ggsave(here("images", "androstenedione_levels_ng_ml.png"), A, width = 3, height = 8)


side_by_side<- P + C + A
ggsave(here("images", "prog_cort_and_levels_controls_ng_ml.png"), side_by_side, width = 8, height = 8)

```



