---
title: "Steroid targeted assay analyses"
output: html_document
date: "2026-04-25"
---

```{r setup, include=FALSE}
#set preference for markdown chunks
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load libraries
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
library(ggplot2)
library(janitor)
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(patchwork)
library(dplyr)
library(ggrepel)
library(magick)
library(caret)
library(glmnet)
library(pROC)


# global publication theme
pub_theme <- theme_minimal(base_size = 11) +
  theme(
    axis.title       = element_text(size = 10),
    axis.text        = element_text(size = 10, colour = "black"),
    axis.line        = element_line(colour = "black", linewidth = 0.3),
    axis.ticks       = element_line(colour = "black", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "grey92", linewidth = 0.3),
    legend.title     = element_text(size = 11),
    legend.text      = element_text(size = 10),
    legend.position  = "bottom",
    strip.text       = element_text(size = 11, face = "plain"),
    plot.tag         = element_text(size = 12, face = "bold"),
    plot.title       = element_text(size = 12, hjust = 0)
  )
theme_set(pub_theme)

# Unified color scheme
risk_cols      <- c(control = "#00BFC4", high = "#F8766D")
risk_labels    <- c(control = "Uncomplicated", high = "Adverse")
timepoint_cols <- c(A = "#a5e35d", B = "#388E3C")
outcome_cols   <- c(Term = "#1cd41c", Preterm = "#e34a0e")
box_fill       <- "#4682b4"
```

# Load data and merge with metadata
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Load metadata
metadata <- read.csv(here("metadata.csv"))
metadata <- clean_names(metadata)

#Load steroids values
hormones <- read.csv((here("hormones_values_final.csv")))
hormones <- clean_names(hormones)

#Merge by sample_id
data <- merge(metadata, hormones, by="sample_id")

#Recode Miscarriage (n=1) as preterm
data <- data %>% mutate (preterm_v_term_1 = case_when(preterm_v_term == "Preterm" ~ "Preterm",
                                                     preterm_v_term == "Miscarriage" ~ "Preterm",
                                                     preterm_v_term == "Term" ~ "Term")) %>%
                                           relocate(preterm_v_term_1, .after = preterm_v_term)

```

# Figure 1
Layout: (A = sample-extraction workflow)
on top, (B = per-subject gestational age at sampling, faceted by Outcome) and
(C = boxplots of P / C / A levels across all samples + median ± SD table) below.
```{r, echo=TRUE, message=FALSE, warning=FALSE}

# Read A from separate Biorender png
extraction_img <- magick::image_read(here("ExtractionWorflow.png"))
panel_A <- patchwork::wrap_elements(
  full = grid::rasterGrob(grDevices::as.raster(extraction_img), interpolate = TRUE)
)

data_panelB <- data %>%
  mutate(
    outcome_grp = factor(
      ifelse(risk_group == "control", "Uncomplicated", "Adverse"),
      levels = c("Uncomplicated", "Adverse")
    ),
    subject = factor(cons_v2)
  )

panel_B <- ggplot(data_panelB,
                  aes(x = gestational_age_final,
                      y = subject,
                      colour = risk_group)) +
  geom_point(size = 2.2, alpha = 0.9) +
  facet_grid(outcome_grp ~ ., scales = "free_y", space = "free_y", switch = "y") +
  scale_colour_manual(values = risk_cols, labels = risk_labels, guide = "none") +
  scale_x_continuous(breaks = seq(12, 24, by = 2)) +
  labs(x = "Gestational Age (weeks)", y = "Subject ID") +
  theme(strip.placement   = "outside",
        strip.text.y.left = element_text(angle = 90, face = "bold"),
        axis.text.y       = element_text(size = 8),
        panel.spacing.y   = unit(8, "pt"))

make_box <- function(yvar, xlab) {
  ggplot(data, aes(x = 1, y = .data[[yvar]])) +
    geom_boxplot(fill = box_fill, linewidth = 0.4) +
    geom_jitter(width = 0.15, height = 0,
                colour = "black", size = 1, alpha = 0.6) +
    scale_x_continuous(breaks = NULL) +
    expand_limits(y = c(0, 20)) +
    scale_y_continuous(breaks = c(0, 5, 10, 15, 20)) +
    coord_cartesian(ylim = c(0, 20)) +
    labs(x = xlab, y = "ng/g") +
    theme(legend.position = "none")
}
prog_box <- make_box("progesterone_ng_g",    "Progesterone")
cort_box <- make_box("cortisol_ng_g",        "Cortisol")
andr_box <- make_box("androstenedione_ng_g", "Androstenedione")

# Collect the shared y-axis title
panel_C_box <- (prog_box + cort_box + andr_box) +
  plot_layout(axis_titles = "collect_y")

# Median & SD table 
fmt_med_sd <- function(x) sprintf("%.2f ± %.2f",
                                  median(x, na.rm = TRUE),
                                  sd(x,     na.rm = TRUE))
sd_tbl <- data.frame(
  Steroid        = c("Progesterone (ng/g)",
                     "Cortisol (ng/g)",
                     "Androstenedione (ng/g)"),
  `Median ± SD`  = c(fmt_med_sd(data$progesterone_ng_g),
                     fmt_med_sd(data$cortisol_ng_g),
                     fmt_med_sd(data$androstenedione_ng_g)),
  check.names = FALSE
)
tt <- gridExtra::ttheme_minimal(
  core    = list(fg_params = list(fontsize = 9),
                 bg_params = list(fill = c("grey97", "white"))),
  colhead = list(fg_params = list(fontsize = 10, fontface = "bold"),
                 bg_params = list(fill = "white"))
)
sd_grob <- gridExtra::tableGrob(sd_tbl, rows = NULL, theme = tt)

panel_C <- wrap_elements(panel_C_box / wrap_elements(full = sd_grob) +
    plot_layout(heights = c(2.0, 1)))

fig1 <- panel_A / (panel_B | panel_C) +
  plot_layout(heights = c(1.0, 1.6), widths = c(1, 1)) +
  plot_annotation(tag_levels = list(c("A", "B", "C"))) &
  theme(plot.tag = element_text(size = 16, face = "bold"))

ggsave(here("images_reviewed", "Figure1.png"), fig1,
       width = 300, height = 240, units = "mm", dpi = 300)
ggsave(here("images_reviewed", "Figure1.svg"), fig1,
       width = 300, height = 240, units = "mm", dpi = 300)

fig1
```

### Basic summary of statistics - all samples included
Summary of basic statistics for progesterone, cortisol and androstenedione levels in all samples. Results saved in "summary_levels_P_C_A.csv". 
```{r}
#Summary statistics progesterone
summary_P<- data %>%
  summarise(
    count = n(),
    min = min(data$progesterone_ng_g, na.rm = TRUE),
    median = median(data$progesterone_ng_g, na.rm = TRUE),
    max = max(data$progesterone_ng_g, na.rm = TRUE),
    sd = sd(data$progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(data$progesterone_ng_g, na.rm = TRUE))

#Summary statistics cortisol
summary_C<- data %>%
  summarise(
    count = n(),
    min = min(data$cortisol_ng_g, na.rm = TRUE),
    median = median(data$cortisol_ng_g, na.rm = TRUE),
    max = max(data$cortisol_ng_g, na.rm = TRUE),
    sd = sd(data$cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(data$cortisol_ng_g, na.rm = TRUE))

#Summary statistics androstenedione
summary_A<- data %>%
  summarise(
    count = n(),
    min = min(data$androstenedione_ng_g, na.rm = TRUE),
    median = median(data$androstenedione_ng_g, na.rm = TRUE),
    max = max(data$androstenedione_ng_g, na.rm = TRUE),
    sd = sd(data$androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(data$androstenedione_ng_g, na.rm = TRUE))

#write.csv(rbind(summary_P, summary_C, summary_A), here("images", "summary_levels_P_C_A.csv"))
summary_P
summary_C
summary_A
```

### Box plots - all samples including
Box plots showing levels of the steroids in all samples available. Image with 3 plots side by side also shown.
```{r}
#box plot of progesterone levels
P <- ggplot(data, aes(x = 1, y = progesterone_ng_g)) +
     geom_boxplot(fill = box_fill, linewidth = 0.4) +
     geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
     scale_x_continuous(breaks = NULL) +
     xlab("Progesterone") + ylab("ng/g") +
     theme(legend.position = "none") +
     coord_cartesian(ylim = c(0, 20))
#ggsave(here("images", "progesterone_levels.png"), P, width = 3, height = 8, dpi = 300)

#box plot of cortisol levels
C <- ggplot(data, aes(x = 1, y = cortisol_ng_g)) +
     geom_boxplot(fill = box_fill, linewidth = 0.4) +
     geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
     scale_x_continuous(breaks = NULL) +
     xlab("Cortisol") + ylab("ng/g") +
     theme(legend.position = "none") +
     coord_cartesian(ylim = c(0, 20))
#ggsave(here("images", "cortisol_levels.png"), C, width = 3, height = 8, dpi = 300)

#box plot of androstenedione levels
A <- ggplot(data, aes(x = 1, y = androstenedione_ng_g)) +
     geom_boxplot(fill = box_fill, linewidth = 0.4) +
     geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
     scale_x_continuous(breaks = NULL) +
     xlab("Androstenedione") + ylab("ng/g") +
     theme(legend.position = "none") +
     coord_cartesian(ylim = c(0, 20))
#ggsave(here("images", "androstenedione_levels.png"), A, width = 3, height = 8, dpi = 300)


#image with 3 plots together
P_C_A_plot <- P + C + A & ylab(NULL)
P_C_A_plot <- wrap_elements(P_C_A_plot) +
  labs(tag = "ng/g") +
  theme(plot.tag = element_text(size = 11, angle = 90),
        plot.tag.position = "left")

#ggsave(here("images", "prog_cort_and_levels.png"), P_C_A_plot, width = 8, height = 8)
P_C_A_plot
```

### Basic summary of statistics - only low-risk samples
Summary of basic statistics for progesterone, cortisol and androstenedione levels in low risk samples.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#filter out high risk pregnancy and term pregnancy
data_low_risk<- data %>% filter(data$risk_group == "control" & data$preterm_v_term_1 == "Term")


summary_P<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    median = median(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    max = max(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    sd = sd(data_low_risk$progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(data_low_risk$progesterone_ng_g, na.rm = TRUE))

summary_C<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    median = median(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    max = max(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    sd = sd(data_low_risk$cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(data_low_risk$cortisol_ng_g, na.rm = TRUE))

summary_A<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    median = median(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    max = max(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    sd = sd(data_low_risk$androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(data_low_risk$androstenedione_ng_g, na.rm = TRUE))

#write.csv(rbind(summary_P, summary_C, summary_A), here("images","summary_levels_P_C_A_controls.csv"))

summary_P
summary_C
summary_A
```

### Box plots - low risk samples
Box plots showing levels of the steroids low risk samples.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#box plot of progesterone levels
P <- ggplot(data_low_risk, aes(x = 1, y = progesterone_ng_g)) +
     geom_boxplot(fill = box_fill, linewidth = 0.4) +
     geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
     scale_x_continuous(breaks = NULL) +
     xlab("Progesterone") + ylab("ng/g") +
     theme(legend.position = "none") +
     coord_cartesian(ylim = c(0, 25))
#ggsave(here("images", "progesterone_levels_LowRisk.png"), P, width = 3, height = 8, dpi = 300)

#box plot of cortisol levels
C <- ggplot(data_low_risk, aes(x = 1, y = cortisol_ng_g)) +
     geom_boxplot(fill = box_fill, linewidth = 0.4) +
     geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
     scale_x_continuous(breaks = NULL) +
     xlab("Cortisol") + ylab("ng/g") +
     theme(legend.position = "none") +
     coord_cartesian(ylim = c(0, 25))
#ggsave(here("images", "cortisol_levels_LowRisk.png"), C, width = 3, height = 8, dpi = 300)

#box plot of androstenedione levels
A <- ggplot(data_low_risk, aes(x = 1, y = androstenedione_ng_g)) +
     geom_boxplot(fill = box_fill, linewidth = 0.4) +
     geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
     scale_x_continuous(breaks = NULL) +
     xlab("Androstenedione") + ylab("ng/g") +
     theme(legend.position = "none") +
     coord_cartesian(ylim = c(0, 25))
#ggsave(here("images", "androstenedione_levels_LowRisk.png"), A, width = 3, height = 8, dpi = 300)


#image with 3 plots together
P_C_A_plot <- P + C + A & ylab(NULL)
P_C_A_plot <- wrap_elements(P_C_A_plot) +
  labs(tag = "ng/g") +
  theme(plot.tag = element_text(size = 11, angle = 90),
        plot.tag.position = "left")

#ggsave(here("images", "prog_cort_and_levels_LowRisk.png"), P_C_A_plot, width = 8, height = 8)
P_C_A_plot
```

## Analysis of steroids relationship with gestational age
### Progesterone
Linear model progesterone ~ gestational age
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#linear model prog vs gest age
model<- lm(data$progesterone_ng_g ~ gestational_age_final , data= data)
s<- summary(model)

R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)
slope <- round(coef(model)[2], 3)
intercept <- round(coef(model)[1], 3)
eq <- paste0("y = ", slope, "x ", ifelse(intercept >= 0, "+ ", "- "), abs(intercept))

# Create label
label_text <- paste0("  r² = ", R2,
                     "\n  ", eq,
                     "\n  p = ", pval)

#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot coloured according to risk factor annotated with stats
prog_scatterplot <- ggplot(data, aes(x = gestational_age_final, y = progesterone_ng_g)) +
  geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction),
            colour = "black", linewidth = 0.5) +
  annotate("text",
           x = -Inf, y = Inf,
           label = label_text, hjust = 0, vjust = 1.2,
           size = 3.2, lineheight = 0.95) +
  labs(colour = "Outcome") +
  xlab("Gestational Age (weeks)") +
  ylab("Progesterone (ng/g)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  scale_colour_manual(values = risk_cols, labels = risk_labels)

#ggsave(here("images_reviewed", "prog_scatterplot_coloured.png"), prog_scatterplot, width = 8, height = 8, dpi = 300)
prog_scatterplot
```

### Cortisol
Linear model cortisol ~ gestational age
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#linear model prog vs gest age
model_cort<- lm(cortisol_ng_g ~ gestational_age_final , data= data)
sc<- summary(model_cort)

R2 <- round(sc$r.squared, 3)
Fstat <- round(sc$fstatistic[1], 3)
df1 <- sc$fstatistic[2]
df2 <- sc$fstatistic[3]
pval <- signif(pf(sc$fstatistic[1], df1, df2, lower.tail = FALSE), 3)
slope <- round(coef(model_cort)[2], 3)
intercept <- round(coef(model_cort)[1], 3)
eq <- paste0("y = ", slope, "x ", ifelse(intercept >= 0, "+ ", "- "), abs(intercept))

# Create label
label_text <- paste0("  r² = ", R2,
                     "\n  ", eq,
                     "\n  p = ", pval)

#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model_cort, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot coloured according to risk factor annotated with stats
cort_scatterplot <- ggplot(data, aes(x = gestational_age_final, y = cortisol_ng_g)) +
  geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction),
            colour = "black", linewidth = 0.5) +
  annotate("text",
           x = -Inf, y = Inf,
           label = label_text, hjust = 0, vjust = 1.2,
           size = 3.2, lineheight = 0.95) +
  labs(colour = "Outcome") +
  xlab("Gestational Age (weeks)") +
  ylab("Cortisol (ng/g)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  scale_colour_manual(values = risk_cols, labels = risk_labels)

#ggsave(here("images_reviewed", "cort_scatterplot_coloured.png"), cort_scatterplot, width = 8, height = 8, dpi = 300)
cort_scatterplot
```

### Androstenedione
Linear model androstenedione ~ gestational age
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#linear model prog vs gest age
model_and <- lm(androstenedione_ng_g ~ gestational_age_final , data= data)
sa <- summary(model_and)

R2 <- round(sa$r.squared, 3)
Fstat <- round(sa$fstatistic[1], 3)
df1 <- sa$fstatistic[2]
df2 <- sa$fstatistic[3]
pval <- signif(pf(sa$fstatistic[1], df1, df2, lower.tail = FALSE), 3)
slope <- round(coef(model_and)[2], 3)
intercept <- round(coef(model_and)[1], 3)
eq <- paste0("y = ", slope, "x ", ifelse(intercept >= 0, "+ ", "- "), abs(intercept))

# Create label
label_text <- paste0("  r² = ", R2,
                     "\n  ", eq,
                     "\n  p = ", pval)


#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction <- predict(model_and, newdata=fake_gest_age)
predicted_frame <-cbind(fake_gest_age, prediction)

#plot coloured according to risk factor
androst_scatterplot <- ggplot(data, aes(x = gestational_age_final, y = androstenedione_ng_g)) +
  geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction),
            colour = "black", linewidth = 0.5) +
  annotate("text",
           x = -Inf, y = Inf,
           label = label_text, hjust = 0, vjust = 1.2,
           size = 3.2, lineheight = 0.95) +
  labs(colour = "Outcome") +
  xlab("Gestational Age (weeks)") +
  ylab("Androstenedione (ng/g)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  scale_colour_manual(values = risk_cols, labels = risk_labels)

#ggsave(here("images_reviewed", "androst_scatterplot_coloured.png"), androst_scatterplot, width = 8, height = 8, dpi = 300)
androst_scatterplot
```

### Figure with the 3 scatterplots
```{r, echo=TRUE, message=FALSE, warning=FALSE}
scatterplots <- prog_scatterplot+ cort_scatterplot+ androst_scatterplot + plot_layout(guides = 'collect')
ggsave(here("images_reviewed", "scatterplots_P_C_A.png"), scatterplots, width = 18, height = 6)
```

## Relationship with gestational age in longitudinal samples
Investigate what happens to hormones levels for individuals were we have longitudinal samples (n=5)
```{r}
#filter only longitudinal data
data_paired<- data %>% filter(data$longitudinal =="yes")
```

### Progesterone
```{r}
#progesterone individuals levels
individuals_prog <- data_paired %>%
  ggplot(aes(gestational_age_final, progesterone_ng_g, colour = timepoint)) +
  geom_line(aes(group = cons_v2), colour = "grey70", linewidth = 0.3) +
  geom_point(size = 2.2) +
  xlab("Gestational Age (weeks)") + ylab("Progesterone (ng/g)") +
  facet_grid(. ~ cons_v2) +
  scale_x_continuous(breaks = c(15, 22)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  scale_colour_manual(values = timepoint_cols, name = "Timepoint") +
  theme(axis.text.x     = element_text(size = 9),
        strip.text      = element_text(size = 9),
        panel.spacing.x = unit(6, "pt"))
#ggsave(here("images_reviewed", "individuals_prog_coloured.png"), individuals_prog, width = 8, height = 8, dpi = 300)

#statistical summary
group_by(data_paired, timepoint) %>%
  summarise(
    count = n(),
    median = median(progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(progesterone_ng_g, na.rm = TRUE))

#box plot
prog_timepoint_A_B <- ggplot(data_paired, aes(x = timepoint, y = as.numeric(progesterone_ng_g))) +
  geom_boxplot(aes(fill = timepoint), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Timepoint") + ylab("Progesterone (ng/g)") +
  scale_fill_manual(values = timepoint_cols) +
  coord_cartesian(ylim = c(0, 20)) +
  theme(legend.position = "none")
#ggsave(here("images_reviewed", "prog_timepoint_A_B_coloured.png"), prog_timepoint_A_B, width = 8, height = 8, dpi = 300)

# wilcoxon pairedtest: A vs B per subject (paired by cons_v2)
wide_prog <- data_paired %>%
  select(cons_v2, timepoint, progesterone_ng_g) %>%
  pivot_wider(names_from = timepoint, values_from = progesterone_ng_g)
result <- wilcox.test(wide_prog$A, wide_prog$B, paired = TRUE)
result$p.value

individuals_prog
prog_timepoint_A_B
``` 

### Cortisol
```{r}
#Cortisol individual levels
individuals_cort <- data_paired %>%
  ggplot(aes(gestational_age_final, cortisol_ng_g, colour = timepoint)) +
  geom_line(aes(group = cons_v2), colour = "grey70", linewidth = 0.3) +
  geom_point(size = 2.2) +
  xlab("Gestational Age (weeks)") + ylab("Cortisol (ng/g)") +
  facet_grid(. ~ cons_v2) +
  scale_x_continuous(breaks = c(15, 22)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  scale_colour_manual(values = timepoint_cols, name = "Timepoint") +
  theme(axis.text.x     = element_text(size = 9),
        strip.text      = element_text(size = 9),
        panel.spacing.x = unit(6, "pt"))
#ggsave(here("images_reviewed", "individuals_cort_coloured.png"), individuals_cort, width = 8, height = 8, dpi = 300)

#statistical summary
group_by(data_paired, timepoint) %>%
  summarise(
    count = n(),
    median = median(cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(cortisol_ng_g, na.rm = TRUE))

#box plot
cort_timepoint_A_B <- ggplot(data_paired, aes(x = timepoint, y = as.numeric(cortisol_ng_g))) +
  geom_boxplot(aes(fill = timepoint), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Timepoint") + ylab("Cortisol (ng/g)") +
  scale_fill_manual(values = timepoint_cols, name = "Timepoint", guide = "none") +
  theme(legend.position = "none")
#ggsave(here("images_reviewed", "cort_timepoint_A_B_coloured.png"), cort_timepoint_A_B, width = 8, height = 8, dpi = 300)

#paired t-test: A vs B per subject (paired by cons_v2)
wide_cort <- data_paired %>%
  select(cons_v2, timepoint, cortisol_ng_g) %>%
  pivot_wider(names_from = timepoint, values_from = cortisol_ng_g)
result <- wilcox.test(wide_cort$A, wide_cort$B, paired = TRUE)
result$p.value

individuals_cort
cort_timepoint_A_B
``` 

### Androstenedione
```{r}
#Androstenedione individual levels
individuals_andr <- data_paired %>%
  ggplot(aes(gestational_age_final, androstenedione_ng_g, colour = timepoint)) +
  geom_line(aes(group = cons_v2), colour = "grey70", linewidth = 0.3) +
  geom_point(size = 2.2) +
  xlab("Gestational Age (weeks)") + ylab("Androstenedione (ng/g)") +
  facet_grid(. ~ cons_v2) +
  scale_x_continuous(breaks = c(15, 22)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  scale_colour_manual(values = timepoint_cols, name = "Timepoint") +
  theme(axis.text.x     = element_text(size = 9),
        strip.text      = element_text(size = 9),
        panel.spacing.x = unit(6, "pt"))
#ggsave(here("images_reviewed", "individuals_andr_coloured.png"), individuals_andr, width = 8, height = 8, dpi = 300)

#statistical summary
group_by(data_paired, timepoint) %>%
  summarise(
    count = n(),
    median = median(androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(androstenedione_ng_g, na.rm = TRUE))

#box plot
andr_timepoint_A_B <- ggplot(data_paired, aes(x = timepoint, y = as.numeric(androstenedione_ng_g))) +
  geom_boxplot(aes(fill = timepoint), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Timepoint") + ylab("Androstenedione (ng/g)") +
  scale_fill_manual(values = timepoint_cols, name = "Timepoint", guide = "none") +
  theme(legend.position = "none")
#ggsave(here("images_reviewed", "andr_timepoint_A_B_coloured.png"), andr_timepoint_A_B, width = 8, height = 8, dpi = 300)

#paired wilcoxon: A vs B per subject (paired by cons_v2)
wide_andr <- data_paired %>%
  select(cons_v2, timepoint, androstenedione_ng_g) %>%
  pivot_wider(names_from = timepoint, values_from = androstenedione_ng_g)
result <- wilcox.test(wide_andr$A, wide_andr$B, paired = TRUE)
result$p.value

individuals_andr
andr_timepoint_A_B
``` 

# Figure 2
```{r}
row_a <- prog_scatterplot   + cort_scatterplot   + androst_scatterplot
row_b <- individuals_prog   + individuals_cort   + individuals_andr
row_c <- prog_timepoint_A_B + cort_timepoint_A_B + andr_timepoint_A_B

#row A: collect Outcome legend + share x-axis title across the row
row_a_grp <- row_a +
  plot_layout(guides = "collect", axis_titles = "collect_x") &
  theme(legend.position = "right")

#rows B+C: collect a single Timepoint legend + share x-axis titles within each row
row_bc_grp <- (row_b + plot_layout(axis_titles = "collect_x")) /
              (row_c + plot_layout(axis_titles = "collect_x")) +
  plot_layout(guides = "collect") &
  guides(fill = "none") &
  theme(legend.position = "right")

#one tag per row (A on top-left, B on row 2 left, C on row 3 left), blanks elsewhere
fig2 <- (row_a_grp / row_bc_grp) +
  plot_layout(heights = c(1, 2)) +
  plot_annotation(tag_levels = list(c("A", "", "", "B", "", "", "C", "", ""))) &
  theme(plot.tag = element_text(size = 16, face = "bold"))

ggsave(here("images_reviewed", "Figure2.png"), fig2,
       width = 300, height = 200, units = "mm", dpi = 300)

ggsave(here("images_reviewed", "Figure2.svg"), fig2, 
       width = 300, height = 200, units = "mm", dpi=300)

fig2
```

# Regression model of cervical lenght and steroids levels
## Progesterone plot
```{r}
data$cl_visit <- as.numeric(as.character(data$cl_visit))

# Remove missing value from miscarriage patient.
data_clean <- data[is.finite(data$cl_visit) & data$cl_visit > 0 & is.finite(data$progesterone_ng_g), ]

# Fit linear model
model_cl <- lm(progesterone_ng_g ~ cl_visit, data = data_clean)

# Summary
s <- summary(model_cl)

# Extract stats
R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)
slope <- round(coef(model_cl)[2], 3)
intercept <- round(coef(model_cl)[1], 3)
eq <- paste0("y = ", slope, "x ", ifelse(intercept >= 0, "+ ", "- "), abs(intercept))

label_text <- paste0("  r² = ", R2,
                     "\n  ", eq,
                     "\n  p = ", pval)

# Prediction frame
fake_cl <- data.frame(cl_visit = seq(min(data_clean$cl_visit),
                                     max(data_clean$cl_visit),
                                     length.out = 100))
prediction <- predict(model_cl, newdata = fake_cl)
predicted_frame <- cbind(fake_cl, prediction)

# Plot
prog_cerv <- ggplot(data_clean, aes(x = cl_visit, y = progesterone_ng_g)) +
  geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = cl_visit, y = prediction),
            colour = "black", linewidth = 0.5) +
  annotate("text",
           x = -Inf, y = Inf,
           label = label_text, hjust = 0, vjust = 1.2,
           size = 3.2, lineheight = 0.95) +
  labs(colour = "Outcome") +
  xlab("Cervical Length (mm)") +
  ylab("Progesterone (ng/g)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  scale_colour_manual(values = risk_cols, labels = risk_labels)
#ggsave(here("images_reviewed", "prog_cerv.png"), prog_cerv, width = 6, height = 6, dpi = 300)
```

## Cortisol plot
```{r}
model_cort <- lm(cortisol_ng_g ~ cl_visit, data = data_clean)
s <- summary(model_cort)

R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)
slope <- round(coef(model_cort)[2], 3)
intercept <- round(coef(model_cort)[1], 3)
eq <- paste0("y = ", slope, "x ", ifelse(intercept >= 0, "+ ", "- "), abs(intercept))

label_text <- paste0("  r² = ", R2,
                     "\n  ", eq,
                     "\n  p = ", pval)

fake_cl <- data.frame(cl_visit = seq(min(data_clean$cl_visit), max(data_clean$cl_visit), length.out = 100))
prediction <- predict(model_cort, newdata = fake_cl)
predicted_frame <- cbind(fake_cl, prediction)

cortisol_cerv <- ggplot(data_clean, aes(x = cl_visit, y = cortisol_ng_g)) +
  geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = cl_visit, y = prediction),
            colour = "black", linewidth = 0.5) +
  annotate("text",
           x = -Inf, y = Inf,
           label = label_text, hjust = 0, vjust = 1.2,
           size = 3.2, lineheight = 0.95) +
  labs(colour = "Outcome") +
  xlab("Cervical Length (mm)") +
  ylab("Cortisol (ng/g)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  scale_colour_manual(values = risk_cols, labels = risk_labels)
#ggsave(here("images_reviewed", "cortisol_cerv.png"), cortisol_cerv, width = 6, height = 6, dpi = 300)

```

## Androstenedione
```{r}
model_and <- lm(androstenedione_ng_g ~ cl_visit, data = data_clean)
s <- summary(model_and)

R2 <- round(s$r.squared, 3)
Fstat <- round(s$fstatistic[1], 3)
df1 <- s$fstatistic[2]
df2 <- s$fstatistic[3]
pval <- signif(pf(s$fstatistic[1], df1, df2, lower.tail = FALSE), 3)
slope <- round(coef(model_and)[2], 3)
intercept <- round(coef(model_and)[1], 3)
eq <- paste0("y = ", slope, "x ", ifelse(intercept >= 0, "+ ", "- "), abs(intercept))

label_text <- paste0("  r² = ", R2,
                     "\n  ", eq,
                     "\n  p = ", pval)

fake_cl <- data.frame(cl_visit = seq(min(data_clean$cl_visit), max(data_clean$cl_visit), length.out = 100))
prediction <- predict(model_and, newdata = fake_cl)
predicted_frame <- cbind(fake_cl, prediction)

andr_cerv <- ggplot(data_clean, aes(x = cl_visit, y = androstenedione_ng_g)) +
  geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = cl_visit, y = prediction),
            colour = "black", linewidth = 0.5) +
  annotate("text",
           x = -Inf, y = Inf,
           label = label_text, hjust = 0, vjust = 1.2,
           size = 3.2, lineheight = 0.95) +
  labs(colour = "Outcome") +
  xlab("Cervical Length (mm)") +
  ylab("Androstenedione (ng/g)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  scale_colour_manual(values = risk_cols, labels = risk_labels)
#ggsave(here("images_reviewed", "andr_cerv.png"), andr_cerv, width = 6, height = 6, dpi = 300)

```


### Figure with the 3 scatterplots
```{r, echo=TRUE, message=FALSE, warning=FALSE}
scatterplots <- prog_cerv + cortisol_cerv + andr_cerv + plot_layout(guides = 'collect')
ggsave(here("images_reviewed", "scatterplots_P_C_A_cerv_lenght.png"), scatterplots,
       width = 18, height = 6, dpi = 300)
```


## Difference in steroids levels in term vs preterm
Analyses performed considering only one sample per patient (select timepoint B for longitudinal ones to balance gestational age of preterm group).
```{r}
#filter data
data_lo <- data %>% filter(data$longitudinal =="yes" & data$timepoint == "B")
data_nolo <- data %>% filter(data$longitudinal =="no")

data_oneeach <- rbind(data_lo, data_nolo)
```

### Progesterone
```{r}
#box plot
boxplot_outcome_prog <- ggplot(data_oneeach, aes(x = preterm_v_term_1, y = as.numeric(progesterone_ng_g))) +
  geom_boxplot(aes(fill = preterm_v_term_1), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Outcome") + ylab("Progesterone (ng/g)") +
  scale_fill_manual(values = outcome_cols) +
  theme(legend.position = "none")
#ggsave(here("images_reviewed", "boxplot_outcome_prog_coloured.png"), boxplot_outcome_prog, width = 6, height = 6, dpi = 300)

#statistical summary
group_by(data_oneeach, preterm_v_term_1) %>%
  summarise(
    count = n(),
    median = median(progesterone_ng_g, na.rm = TRUE),
    mean=mean(progesterone_ng_g, na.rm = TRUE),
    IQR = IQR(progesterone_ng_g, na.rm = TRUE),
    SD = sd(progesterone_ng_g))

#wilcoxon test
res <- wilcox.test(progesterone_ng_g ~ preterm_v_term_1, data = data_oneeach,
                   exact = FALSE)
res
boxplot_outcome_prog
```

### Cortisol
```{r}
#box plot
boxplot_outcome_cort <- ggplot(data_oneeach, aes(x = preterm_v_term_1, y = as.numeric(cortisol_ng_g))) +
  geom_boxplot(aes(fill = preterm_v_term_1), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Outcome") + ylab("Cortisol (ng/g)") +
  scale_fill_manual(values = outcome_cols) +
  theme(legend.position = "none")
#ggsave(here("images_reviewed", "boxplot_outcome_cort_coloured.png"), boxplot_outcome_cort, width = 6, height = 6, dpi = 300)

#statistical summary
group_by(data_oneeach, preterm_v_term_1) %>%
  summarise(
    count = n(),
    median = median(cortisol_ng_g, na.rm = TRUE),
    mean=mean(cortisol_ng_g, na.rm = TRUE),
    IQR = IQR(cortisol_ng_g, na.rm = TRUE),
    SD = sd(cortisol_ng_g))

#wilcoxon test
res <- wilcox.test(cortisol_ng_g ~ preterm_v_term_1, data = data_oneeach,
                   exact = FALSE)
res
boxplot_outcome_cort
```

### Androstenedione
```{r}
#boxplot
boxplot_outcome_andr <- ggplot(data_oneeach, aes(x = preterm_v_term_1, y = as.numeric(androstenedione_ng_g))) +
  geom_boxplot(aes(fill = preterm_v_term_1), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Outcome") + ylab("Androstenedione (ng/g)") +
  scale_fill_manual(values = outcome_cols) +
  theme(legend.position = "none")
ggsave(here("images_reviewed", "boxplot_outcome_andr_coloured.png"), boxplot_outcome_andr,
       width = 6, height = 6, dpi = 300)

#statistical summary
group_by(data_oneeach, preterm_v_term_1) %>%
  summarise(
    count = n(),
    median = median(androstenedione_ng_g, na.rm = TRUE),
    IQR = IQR(androstenedione_ng_g, na.rm = TRUE),
    SD = sd(androstenedione_ng_g))

#wilcoxon test
res <- wilcox.test(androstenedione_ng_g ~ preterm_v_term_1, data = data_oneeach,
                   exact = FALSE)
res
boxplot_outcome_andr
```

### save image with 3 boxplots
```{r}
boxplot_outcome_all <- boxplot_outcome_prog + boxplot_outcome_cort + boxplot_outcome_andr + plot_layout(guides = 'collect')
ggsave(here("images_reviewed", "boxplot_outcome_P_C_A.png"), boxplot_outcome_all,
       width = 18, height = 6, dpi = 300)
```

```{r}
# Combine rows
combined_figure <- scatterplots / boxplot_outcome_all +
    plot_layout(heights = c(1, 1.2))  # adjust spacing

# Save
ggsave(here("images_reviewed", "combined_scatter_boxplots_panelAB.png"),
       combined_figure, width = 18, height = 12, dpi = 300)
```

# Figure 3
```{r, echo=TRUE, message=FALSE, warning=FALSE}
row_a_f3 <- (prog_cerv + cortisol_cerv + andr_cerv) +
  plot_layout(guides = "collect", axis_titles = "collect_x") &
  theme(legend.position = "right")

row_b_f3 <- (boxplot_outcome_prog + boxplot_outcome_cort + boxplot_outcome_andr) +
  plot_layout(guides = "collect", axis_titles = "collect_x") &
  guides(fill = "none") &
  theme(legend.position = "none")

fig3 <- (row_a_f3 / row_b_f3) +
  plot_layout(heights = c(1, 1)) +
  plot_annotation(tag_levels = list(c("A", "", "", "B", "", ""))) &
  theme(plot.tag = element_text(size = 16, face = "bold"))

ggsave(here("images_reviewed", "Figure3.png"), fig3,
       width = 300, height = 200, units = "mm", dpi = 300)
ggsave(here("images_reviewed", "Figure3.svg"), fig3,
       width = 300, height = 200, units = "mm", dpi = 300)

fig3
```


```{r}
# Convert weeks + days into decimal weeks
data$gestational_age_delivery_final <- data$gestation_at_delivery + data$gestation_at_delivery_days / 7

# Check
head(data[, c("gestation_at_delivery", "gestation_at_delivery_days", "gestational_age_delivery_final")])
```

```{r}
# Progesterone
model_prog_gad <- lm(progesterone_ng_g ~ gestational_age_delivery_final, data = data)
summary(model_prog_gad)

# Cortisol
model_cort_gad <- lm(cortisol_ng_g ~ gestational_age_delivery_final, data = data)
summary(model_cort_gad)

# Androstenedione
model_and_gad <- lm(androstenedione_ng_g ~ gestational_age_delivery_final, data = data)
summary(model_and_gad)

```

```{r}
# Function to extract R², F-statistic, p-value
get_stats <- function(model){
  s <- summary(model)
  R2 <- round(s$r.squared, 3)
  Fstat <- round(s$fstatistic[1], 3)
  df1 <- s$fstatistic[2]
  df2 <- s$fstatistic[3]
  pval <- signif(pf(Fstat, df1, df2, lower.tail = FALSE), 3)
  slope <- round(coef(model)[2], 3)
  intercept <- round(coef(model)[1], 3)
  return(list(R2 = R2, Fstat = Fstat, df1 = df1, df2 = df2, pval = pval,
              slope = slope, intercept = intercept))
}

stats_prog <- get_stats(model_prog_gad)
stats_cort <- get_stats(model_cort_gad)
stats_and <- get_stats(model_and_gad)
```


# Supplementary Figure S4 
```{r}
# Function to create plot
plot_steroid <- function(data, yvar, model, stats, ylab){

  # Prediction frame
  fake_x <- data.frame(gestational_age_delivery_final = seq(min(data$gestational_age_delivery_final, na.rm = TRUE),
                                                      max(data$gestational_age_delivery_final, na.rm = TRUE),
                                                      length.out = 100))
  prediction <- predict(model, newdata = fake_x)
  predicted_frame <- cbind(fake_x, prediction)

  # Label text
  eq <- paste0("y = ", stats$slope, "x ",
               ifelse(stats$intercept >= 0, "+ ", "- "), abs(stats$intercept))
  label_text <- paste0("  r² = ", stats$R2,
                       "\n  ", eq,
                       "\n  p = ", stats$pval)

  # Plot
  ggplot(data, aes(x = gestational_age_delivery_final, y = !!sym(yvar))) +
    geom_point(aes(colour = risk_group), size = 2.2, alpha = 0.85) +
    geom_line(data = predicted_frame, aes(x = gestational_age_delivery_final, y = prediction),
              colour = "black", linetype = "dashed", linewidth = 0.5) +
    annotate("text",
             x = -Inf, y = Inf,
             label = label_text, hjust = 0, vjust = 1.2,
             size = 3.2, lineheight = 0.95) +
    scale_colour_manual(values = risk_cols, labels = risk_labels) +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
    labs(x = "Gestational Age at Delivery (weeks)", y = ylab, colour = "Outcome")
}

# Create plots
prog_plot <- plot_steroid(data, "progesterone_ng_g", model_prog_gad, stats_prog, "Progesterone (ng/g)")
cort_plot <- plot_steroid(data, "cortisol_ng_g", model_cort_gad, stats_cort, "Cortisol (ng/g)")
and_plot  <- plot_steroid(data, "androstenedione_ng_g", model_and_gad, stats_and, "Androstenedione (ng/g)")

# Display example
prog_plot
cort_plot
and_plot


# Combine plots in a single row — match Figure 2/3 standard:
# shared bottom x-axis title, collected legend, A/B/C panel tags, mm units, 300 dpi, PNG + SVG.
figS4 <- (prog_plot + cort_plot + and_plot) +
  plot_layout(guides = "collect", axis_titles = "collect_x") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 16, face = "bold"))

ggsave(here("images_reviewed", "steroids_vs_gest_age_delivery_noF.png"),
       figS4, width = 300, height = 120, units = "mm", dpi = 300)
ggsave(here("images_reviewed", "FigureS4.png"), figS4,
       width = 300, height = 120, units = "mm", dpi = 300)
ggsave(here("images_reviewed", "FigureS4.svg"), figS4,
       width = 300, height = 120, units = "mm", dpi = 300)

figS4
```


### check if a specific patient has anomalous levels of the steroids
```{r}
# Example for progesterone
patient_sample <- subset(data, sample_id == "2V1592_23+5")

patient_sample

controls <- subset(data, risk_group == "control")

range_prog_ctrl <- range(controls$progesterone_ng_g, na.rm = TRUE)
range_cort_ctrl <- range(controls$cortisol_ng_g, na.rm = TRUE)
range_andr_ctrl <- range(controls$androstenedione_ng_g, na.rm = TRUE)

range_prog_ctrl
range_cort_ctrl
range_andr_ctrl
# Check if within range
prog_in_range <- patient_sample$progesterone_ng_g >= range_prog_ctrl[1] & patient_sample$progesterone_ng_g <= range_prog_ctrl[2]
cort_in_range <- patient_sample$cortisol_ng_g >= range_cort_ctrl[1] & patient_sample$cortisol_ng_g <= range_cort_ctrl[2]
andr_in_range <- patient_sample$androstenedione_ng_g >= range_andr_ctrl[1] & patient_sample$androstenedione_ng_g <= range_andr_ctrl[2]

# Print results
cat("Progesterone within control range:", prog_in_range, "\n")
cat("Cortisol within control range:", cort_in_range, "\n")
cat("Androstenedione within control range:", andr_in_range, "\n")

```

### Multivariate Regression
```{r}
data_clean <- data[is.finite(data$cl_visit) & is.finite(data$progesterone_ng_g), ]

# Convert risk_group to numeric 0/1
data_clean$risk_bin <- ifelse(data_clean$risk_group == "high", 1, 0)

# Run logistic regression
model_multi <- glm(risk_bin ~ progesterone_ng_g + cortisol_ng_g + androstenedione_ng_g, 
                   data = data_clean, family = binomial)
summary(model_multi)
```


```{r}
#adjust for maternal confunders

model_multi_cov <- glm(risk_bin ~ progesterone_ng_g + cortisol_ng_g + androstenedione_ng_g +
                                      age + bmi + parity, data = data_clean, family = binomial)
summary(model_multi_cov)

```

### AUC for multivariate logistic model

```{r}
# Prepare modeling dataset
data_model <- data_clean %>%
  select(risk_group,
         progesterone_ng_g,
         cortisol_ng_g,
         androstenedione_ng_g,
         age, bmi, parity) %>%
  na.omit()

# Ensure outcome is a factor with correct ordering
data_model$risk_group <- factor(
  data_model$risk_group,
  levels = c("control", "high")
)

# Define 3-fold cross-validation
ctrl <- trainControl(
  method = "cv",
  number = 3,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = TRUE
)

# Train logistic regression via caret (glmnet with no regularization)
set.seed(123)
model_cv <- train(
  risk_group ~ progesterone_ng_g +
               cortisol_ng_g +
               androstenedione_ng_g +
               age + bmi + parity,
  data = data_model,
  method = "glmnet",
  metric = "ROC",
  trControl = ctrl,
  tuneGrid = expand.grid(
    alpha = 0,
    lambda = 0
  ),
  family = "binomial"
)

# Extract cross-validated ROC AUC
cv_auc <- max(model_cv$results$ROC)
print(paste("3-fold cross-validated ROC AUC:", round(cv_auc, 3)))

# Plot ROC curve (ggplot version via plotROC)
library(plotROC)

roc_df <- data.frame(
  d = ifelse(model_cv$pred$obs == "high", 1, 0),
  m = model_cv$pred$high
)

roc_plot <- ggplot(roc_df, aes(d = d, m = m)) +
  geom_roc(n.cuts = 0, color = "firebrick", size = 0.6) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  coord_equal() +
  labs(x = "1 - Specificity",
       y = "Sensitivity") +
  annotate("text", x = 0.65, y = 0.10,
           label = paste0("AUC = ", round(cv_auc, 3)),
           size = 3.5)

roc_plot

ggsave(here("images_reviewed", "cross_validated_ROC.png"),
       roc_plot, width = 6, height = 6, dpi = 300)

```


```{r}
pca_res <- prcomp(data_clean[, c("progesterone_ng_g","cortisol_ng_g","androstenedione_ng_g")], scale = TRUE)
scores <- as.data.frame(pca_res$x)
scores$risk_group <- data_clean$risk_group

#variance explained per PC (in %), used for axis labels
var_exp <- summary(pca_res)$importance[2, ] * 100

pca_plot <- ggplot(scores, aes(x = PC1, y = PC2, colour = risk_group)) +
  geom_point(size = 2.2, alpha = 0.85) +
  scale_colour_manual(values = risk_cols, labels = risk_labels) +
  labs(x      = paste0("PC1 (", round(var_exp[1], 1), "%)"),
       y      = paste0("PC2 (", round(var_exp[2], 1), "%)"),
       colour = "Outcome")

ggsave(here("images_reviewed", "PCA_multivariate_adjusted.png"),
       pca_plot, width = 8, height = 6, dpi = 300)


```


```{r}
# Proportion of variance explained
pca_var <- summary(pca_res)$importance[2, ]  # 2nd row = Proportion of Variance
pca_var

# Loadings
pca_res$rotation

```

```{r}
loadings <- as.data.frame(pca_res$rotation[, 1:2])
loadings$analyte <- dplyr::recode(
  rownames(pca_res$rotation),
  progesterone_ng_g    = "Progesterone",
  cortisol_ng_g        = "Cortisol",
  androstenedione_ng_g = "Androstenedione")

# Scaling factor
arrow_scale <- min(
  diff(range(scores$PC1)),
  diff(range(scores$PC2))) * 0.7

loadings$PC1 <- loadings$PC1 * arrow_scale
loadings$PC2 <- loadings$PC2 * arrow_scale


pca_plot_loadings <- ggplot(scores, aes(x = PC1, y = PC2, colour = risk_group)) +
  geom_point(size = 2.2, alpha = 0.85) +

  # Loadings arrows (dashed)
  geom_segment(
    data = loadings,
    aes(x = 0, y = 0, xend = PC1, yend = PC2),
    arrow = arrow(length = unit(0.2, "cm")),
    inherit.aes = FALSE,
    colour = "black", linewidth = 0.4, linetype = "dashed") +

  # Loadings labels with ggrepel to avoid overlap with points / arrows
  ggrepel::geom_text_repel(
    data = loadings,
    aes(x = PC1, y = PC2, label = analyte),
    inherit.aes    = FALSE,
    size           = 3.2,
    seed           = 1,
    box.padding    = 0.5,
    point.padding  = 0.3,
    segment.colour = "grey50",
    segment.size   = 0.3,
    min.segment.length = 0) +

  scale_colour_manual(values = risk_cols, labels = risk_labels) +

  labs(x      = paste0("PC1 (", round(var_exp[1], 1), "%)"),
       y      = paste0("PC2 (", round(var_exp[2], 1), "%)"),
       colour = "Outcome")

ggsave(here("images_reviewed", "PCA_multivariate_loadings.png"),
       pca_plot_loadings, width = 8, height = 6, dpi = 300)

```

# Supplementary Figure 5
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Build coefficients table from the covariate-adjusted logistic model.
coef_mat <- coef(summary(model_multi_cov))
pretty_names <- c(
  "(Intercept)"           = "(Intercept)",
  "progesterone_ng_g"     = "Progesterone (ng/g)",
  "cortisol_ng_g"         = "Cortisol (ng/g)",
  "androstenedione_ng_g"  = "Androstenedione (ng/g)",
  "age"                   = "Maternal age (years)",
  "bmi"                   = "BMI (kg/m²)",
  "parity"                = "Parity"
)
coef_tbl <- data.frame(
  Predictor      = pretty_names[rownames(coef_mat)],
  `Estimate (β)` = formatC(coef_mat[, "Estimate"],   format = "f", digits = 4),
  `Std. Error`   = formatC(coef_mat[, "Std. Error"], format = "f", digits = 3),
  `z value`      = formatC(coef_mat[, "z value"],   format = "f", digits = 2),
  `p-value`      = formatC(coef_mat[, "Pr(>|z|)"],  format = "f", digits = 3),
  check.names = FALSE,
  row.names   = NULL
)
# Drop intercept row from the displayed table.
coef_tbl <- coef_tbl[coef_tbl$Predictor != "(Intercept)", ]

# Bold significant p-values (< 0.05).
sig_idx <- which(as.numeric(coef_tbl$`p-value`) < 0.05)
fg_face <- matrix("plain", nrow = nrow(coef_tbl), ncol = ncol(coef_tbl))
if (length(sig_idx)) fg_face[sig_idx, ncol(coef_tbl)] <- "bold"

tt <- gridExtra::ttheme_minimal(
  core    = list(fg_params = list(fontsize = 10, fontface = fg_face),
                 bg_params = list(fill = c("grey97", "white"))),
  colhead = list(fg_params = list(fontsize = 11, fontface = "bold"),
                 bg_params = list(fill = "white"))
)
coef_grob <- gridExtra::tableGrob(coef_tbl, rows = NULL, theme = tt)

top_row <- (pca_plot_loadings + roc_plot) +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")

figS5 <- (top_row / wrap_elements(full = coef_grob)) +
  plot_layout(heights = c(1.4, 1)) +
  plot_annotation(tag_levels = list(c("A", "B", "C"))) &
  theme(plot.tag = element_text(size = 16, face = "bold"))

ggsave(here("images_reviewed", "FigureS5.png"), figS5,
       width = 300, height = 220, units = "mm", dpi = 300)
ggsave(here("images_reviewed", "FigureS5.svg"), figS5,
       width = 300, height = 220, units = "mm", dpi = 300)

figS5
```


## Variation of CVF fluid weight collected with gestational age
Analyse if there is a significant difference in the weight of the CVF collected according to the gestational age
```{r}
#linear model weight vs gestational age
model<- lm(weight_g ~ gestational_age_final , data= data)
summary(model)

#create dataframe with prediction and fake gestational age (I added some manually so I could do cbind with the values)
fake_gest_age <- data.frame(gestational_age_final = c(seq(12.5, 25.5, by = 0.005)))
prediction<- predict(model, newdata=fake_gest_age)
predicted_frame<-cbind(fake_gest_age, prediction)

#plot
prog_scatterplot_all <- ggplot(data, aes(x = gestational_age_final, y = weight_g)) +
  geom_point(size = 2.2, alpha = 0.85) +
  geom_line(data = predicted_frame, aes(x = gestational_age_final, y = prediction),
            colour = "black", linewidth = 0.5) +
  xlab("Gestational Age (weeks)") + ylab("CVF weight (g)")

ggsave(here("images", "weight_gestational_age.png"), prog_scatterplot_all,
       width = 8, height = 8, dpi = 300)

prog_scatterplot_all
```


## Different weight for longitudinal sample timepoint A and B
```{r}

data_total_paired<- data %>% filter(data$longitudinal =="yes")

weight_timepoint_A_B <- ggplot(data_total_paired, aes(x = timepoint, y = weight_g)) +
  geom_boxplot(aes(fill = timepoint), alpha = 0.8, outlier.shape = NA, linewidth = 0.4) +
  xlab("Timepoint") + ylab("CVF weight (g)") +
  scale_fill_manual(values = timepoint_cols) +
  theme(legend.position = "none")
ggsave(here("images", "weight_timepoint_A_B.png"), weight_timepoint_A_B,
       width = 8, height = 8, dpi = 300)

#paired Wilcoxon: A vs B per subject (paired by cons_v2)
wide_weight <- data_total_paired %>%
  select(cons_v2, timepoint, weight_g) %>%
  pivot_wider(names_from = timepoint, values_from = weight_g)
res_a <- wilcox.test(wide_weight$A, wide_weight$B,
                     exact = FALSE, paired = TRUE, alternative = "two.sided")
res_a$p.value
```


## Analysis with values expressed in ng/ml
Repeat the basic analysis with the value recalculated in unit ng/ml
```{r}

hormones_ng_ml<- read.csv(here("hormones_ng_ml.csv"))

hormones_ng_ml<- clean_names(hormones_ng_ml)

data_ng_ml<-merge(metadata, hormones_ng_ml, by= "sample_id")

data_ng_ml<- data_ng_ml %>% mutate (preterm_v_term_1 = case_when(preterm_v_term == "Preterm" ~ "Preterm",
                                                     preterm_v_term == "Miscarriage" ~ "Preterm",
                                                     preterm_v_term == "Term" ~ "Term")) %>% relocate(preterm_v_term_1, .after = preterm_v_term)
```


### measurment of steroids only in low risk samples
```{r}

data_low_risk<- data_ng_ml %>% filter(data_ng_ml$risk_group == "control" & data_ng_ml$preterm_v_term_1 == "Term")

summary_P<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    median = median(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    max = max(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    sd = sd(data_low_risk$progesterone_ng_ml, na.rm = TRUE),
    IQR = IQR(data_low_risk$progesterone_ng_ml, na.rm = TRUE))

summary_C<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    median = median(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    max = max(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    sd = sd(data_low_risk$cortisol_ng_ml, na.rm = TRUE),
    IQR = IQR(data_low_risk$cortisol_ng_ml, na.rm = TRUE))

summary_A<- data_low_risk %>%
  summarise(
    count = n(),
    min = min(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    median = median(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    max = max(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    sd = sd(data_low_risk$androstenedione_ng_ml, na.rm = TRUE),
    IQR = IQR(data_low_risk$androstenedione_ng_ml, na.rm = TRUE))

write.csv(rbind(summary_P, summary_C, summary_A), here("images","summary_levels_NG_ML_controls.csv"))



P <- ggplot(data_low_risk, aes(x = 1, y = progesterone_ng_ml)) +
  geom_boxplot(fill = box_fill, linewidth = 0.4) +
  geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
  scale_x_continuous(breaks = NULL) +
  xlab("Progesterone") + ylab("ng/ml") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 10))
ggsave(here("images", "progesterone_levels_ng_ml.png"), P,
       width = 3, height = 8, dpi = 300)

C <- ggplot(data_low_risk, aes(x = 1, y = cortisol_ng_ml)) +
  geom_boxplot(fill = box_fill, linewidth = 0.4) +
  geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
  scale_x_continuous(breaks = NULL) +
  xlab("Cortisol") + ylab("ng/ml") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 10))
ggsave(here("images", "cortisol_levels_ng_ml.png"), C,
       width = 3, height = 8, dpi = 300)

A <- ggplot(data_low_risk, aes(x = 1, y = androstenedione_ng_ml)) +
  geom_boxplot(fill = box_fill, linewidth = 0.4) +
  geom_jitter(width = 0.15, height = 0, colour = "black", size = 1, alpha = 0.6) +
  scale_x_continuous(breaks = NULL) +
  xlab("Androstenedione") + ylab("ng/ml") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 10))
ggsave(here("images", "androstenedione_levels_ng_ml.png"), A,
       width = 3, height = 8, dpi = 300)


side_by_side <- P + C + A
ggsave(here("images", "prog_cort_and_levels_controls_ng_ml.png"), side_by_side,
       width = 8, height = 8, dpi = 300)

```



